Influence of Connectivity on Dung Beetle Communities

### 2025-06-06 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
last methods and made figures

- any rare species? 
- any invasive / exotic species?

When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches. Statistical analysis of abundance relating to abundance focused on the six most abundant species. When looking at counts for our most abundant species, results were variable but overall follow a similar pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch. One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.

### _QUESTION FOR EMILIO_ <-------
And then here I want to talk about model M3 in 04_GLMM but this one is tricky there is a lot going on when looking at the model summary what should be the important info to report?

FROM EB: THE BIG BICTURE - sometimes the big picture is a clear effect (abundance of all species was higher in the winged patches). SOmetimes it is a split result (eg abundance of species a higher in winged, abund of species b higher in connected). And sometimes it is messy. that looks like the case here, so say that in thesis language: "the results were highly variable, with some species higher in A (eg, spec 1, 2) whiole others were higher in b (sp 3,4). For some species there was no effect of patch type on abundance (sp 5). Complete results are in table X.  in other words in some cases "a lot going on" IS the message.

EB: there you go! That's how its done.

@@

### Another _QUESTION FOR EMILIO_ <-------
How best can I expand upon this more and same question for hill numbers, maybe means?

Differences in species richness showed no significance across patch types. The total number of species captured in each treatment varied by sampling block, however species totals generally ranged from 8 to 13 species. The patch with the most species captured was in the matrix at block 53n while the fewest were captured at the rectangle patch in block 8 (sp_richness.png).

FROM EB: THIS IS WHAT YOU NEED TO SAY IN THE RESULTS. you can add a little. ("The hill numbers ranged from X in habitat Y to X in habitat z. However, there was no significant...") You expand on what it *means* in the discussion. Here just report the facts.

@@

Values for Hill numbers also showed no significance between patch types. Modeled values for both Shannon and Simpson's diversity had relatively high variance but averaged out at approximately qD = 5.29 and qD = 4.07 respectively.

>>>>

When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches. Statistical analysis of abundance focused on the six most abundant species. When looking at counts for our most abundant species, results were variable but overall follow a consistent pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch. One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types.

@@

The number of species per patch varied from N = 8 (rectangle patch in block 8) to N = 13 (matrix patch in block 53N). Species richness was relatively even across patch types with matrix and connected patches averaging out to approximately N = 11 species and rectangle and winged patches averaging to approximately N = 10 species. Overall there were no significant effects, however species richness patterns were variable between sampling blocks.

@@

Values for Hill numbers showed no patterns for shifts in species composition by patch type. Data for Shannon diversity showed no differences in the distribution of rare species between patch types. On average connected patches had a value of D = 5.29, while mean estimates for all other patches remained close with +0.03% for rectangle, -0.02% for matrix, and -0.06% for winged patches. There was no significance between patch types but data was highly variable between patch types and sampling blocks.

Values for Simpson Diversity also showed no differences but instead for more abundant species composing the dung beetle assemblages. Simpson diversity had a mean value of D = 4.07 in the connected patch, also with other values in close proximity with matrix +0.02%, rectangle +0.04%, and winged at -0.1%. There was no significant difference between patch types, however this data was also highly variable between patch type and sampling blocks.
